GPU-Beschleunigung der Material Point Method

### 2019-01-21 mscthesis.tex
sorting

\newenvironment{code}{\captionsetup{type=algorithm}}{}
 \caption{P2G-Transfers of one million randomly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_random}
\begin{document}
\author{Fabian Meyer}
\title{GPU-Beschleunigung der Material Point Method}
\zweitgutachter{Bastian Krayer, M.Sc.}
\zweitgutachterInfo{(Institut f{\"u}r Computervisualistik, AG Computergraphik)}
% Umschalten der Sprache (fuer englische Rubrikbezeichnungen etc.)

@@

\subsection{Counting Sort \& Stream Compaction}\label{sec:sort}
Sorting on the GPU can dramatically increase workload performance of subsequent steps. If they can profit from following things:
\begin{enumerate}
  \item Accesses can now be handled in a coalesced fashion. \label{it:sort_1}
  \item Data can now be reused due to L2-Cache and shared memory. \label{it:sort_2}
\end{enumerate}
This is the case for algorithms which require neighbor communication.
Supported is either a full deep copy of all particle variables. A deep copy profits from item \ref{it:sort_1} and \ref{it:sort_2}, but takes the hit in performance of uncoalesced accesses for all variables. Or alternatively an indexing to access an existing particle buffer in the specified sorted manner. Indexing only profits from item \ref{it:sort_2} and subsequent shaders will be uncoalesced.

@@

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Method         &  $\Delta t_c$ & Speedup &VRAM              & SM              & Sel. Warp-Stall Reas.\\\hline
    Interleaved    & 305                   & -       &23.0\%          & \textbf{60.9}\%  & S. Scoreb.(17.2\%)\\\hline
    Sequential     & 141                   & 2.16x   &\textbf{49.8}\%   & 37.1\%         & \textbf{S. Scoreb.(2.0}\%) \\\hline
    Seq. (2x)      & 100                   & 3.05x   &69.5\%            & 26.2\%         & L. Scoreb.(80.1\%)\\\hline
    Seq. (128x)    & 98                    & 3.1x    &\textbf{72.9}\%   & 16.9\%         & L. Scoreb.(84.4\%)\\\hline
    Seq. (256x)    & 101                   & 3.0x    &66.4\%            & 14.6\%         & L. Scoreb.(76.9\%)\\
    \hline
  \end{tabular}
  \caption{Optimization of one parallel reduction dispatch on $1024\times1024$ \texttt{vec4} with \texttt{unary\_op(vec4 x)=length(x)} with block\_size = 1024. $\Delta t_c$ is in $\mu s$ Note that the interleaved addressing in this example intentionally commits bank conflicts to show performance impacts. The number in parentheses states the number of additional added sequential elements. Reduced is by block\_size$\times$ sequential elements. Thus they do more work, while also performing better!}\label{tab:map_reduce}
\end{table}

@@

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Interleaved       &  $\Delta t_c$         & Speedup &VRAM              & SM      & L2     & Sel. Stall Reas.\\\hline
    Conflicts           & 748                   & -       &17.5\%            & 51.7\%  & 8.2\%  & S. Scoreb.(6.3\%)\\\hline
    No confl.           & 571                   & 1.31x   &25.7\%            & 50.1\%  & 5.4\%  & \textbf{S. Scoreb.(2.6\%)}\\\hline
    part. confl.        & 546                   & 1.36x   &23.5\%            & 50.8\%  & 6.0\%  & \textbf{S. Scoreb.(3.2\%)} \\\hline\hline
    Seq. (2x)           & 323                   & 2.31x   &41.7\%            & \textbf{43.0}\%  & 17.2\% & S. Scoreb.(32.0\%)\\\hline
    unrolled            & 258                   & 2.90x   &\textbf{53.8}\%   & 42.2\%  & \textbf{22.0}\% & S. Scoreb.(32.0\%)\\\hline\hline
    Seq. (4x)           & 311                   & 2.40x   &56.8\%            & 22.7\%  & 37.9\% & L. Scoreb.(43.3\%)\\\hline
    unrolled            & 297                   & 2.52x   &61.1\%            & 21.2\%  & 40.6\% & L. Scoreb.(49.9\%)\\
    \hline
  \end{tabular}
  \caption{Optimization of one scan dispatch on $4 \times 1024\times1024$ integers with \texttt{binary\_op(x,y) := x+y}. $\Delta t_c$ in $\mu s$, block size is 1024. The number in parentheses states the number of additionally loaded sequential elements. Segment of each scan is block\_size$\times$ sequential elements. Higher reduction factor corresponds to more work being done.}\label{tab:scan}
\end{table}
The second optimization is giving these algorithms more sequential work. This will lighten the SM throughput and increase VRAM throughput as can be seen in both cases. Note, that each thread in table \ref{tab:map_reduce} and \ref{tab:scan} does the same amount of parallel work; the sequential work is additionally being done. However due to the additional sequential work parallelization a block level goes down by the same factor. As mentioned in \cite{NVIDIA:PARALLEL_REDUCTION} this balances the relation of I/O latency to work better and thus improves performance.

@@

For the scan this is a little bit harder as elements need to be stored in registers. The effects can be seen in table \ref{tab:scan}. The segment each scan can handle is by far not as extensive as in parallel reduction. But the speedup achieved from a partially conflict free to an implementation with sequential work is in contrast much higher. Since each thread requires the nearest by elements, L2-cache hits increase rapidly as does the L2 utilization.Thus the scoreboard stall reasons rise again.

\textbf{Counting Sort:}
\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |r|}    \hline
    Ordering            &  $\Delta t_c (\mu s)$ & Speedup &VRAM              & SM      & L2     & L2-Hit \\\hline
    Random              &  1,516                & -       &25.0\%            & 3.4\%   & 9.1\%  & 10.8\% \\\hline
    Deep sort           &    218                & 6.95x   &75.3\%            & 24.4\%  & 35.0\% & 37.8\% \\
    \hline
 \end{tabular}
\caption{Order dependency of binning of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid.}\label{table:bin}
\end{table}

\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r| }    \hline
    Best Case    &  $\Delta t_c (\mu s)$&VRAM  & Top Stall Reas.    & Read   & Write  \\\hline
    Index        &  1,516       &25.6\%& L. Scoreb.(78.7\%) & 16.30\%& 9.30\% \\\hline
    Full(8 vec4) &  1,731       &75.3\%& L. Scoreb.(92.2\%) & 38.60\%& 34.60\%\\
    \hline
    \hline
    Worst Case             &  $\Delta t_c(\mu s)$         &VRAM              & Top Stall Reas.    & Read   & Write\\\hline
    Index                  &  1,516                &26.0\%            & L. Scoreb.(73.5\%) & 16.10\%& 9.40\%\\\hline
    Full(8 vec4)           &  9,847                &24.0\%            & L. Scoreb.(82.2\%) & 9.90\% & 14.0\%\\
    \hline
 \end{tabular}
\caption{Order dependency of binning of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid. Best case is already sorted, worst case is unsorted and still fully random.}\label{table:bin}
\end{table}

\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |r|r|r|}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2    &SM     \\\hline
    G2P global             &   38,318             &-               &24.3\%&34.4\% & 9.6\% \\\hline
    sorted                 &    2,991		  &12.81x          & 7.5\%&52.3\% &67.1\% \\\hline
    G2P-Shared             &    9,665             & 3.96x          & 3.4\%& 2.5\% &65.0\% \\\hline
 \end{tabular}
 \caption{G2P-Transfers of a $128\times 128\times128$ grid back to one million randomly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$. Block size is (8,4,4).}\label{table:p2g_uniform4}
\end{table}

VRAM-L.Scoreboard 79.9\%
Wait No Instructions
\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM    \\\hline
    G2P global             &   10,316             &-               &2.3\%&34.4\%&32.4\% \\\hline
    sorted                 &    2,629		  &3.92x           &4.1\%&9.4\%&66.9\%  \\\hline
    G2P-Shared             &    2,232             &4.62x           &7.3\%&4.1\% &63.6\% \\\hline
 \end{tabular}

 \caption{G2P-Transfers of a $128\times 128\times128$ grid back to one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$. They form a rotated (unsorted) cube with four particles per cell. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_uniform4}
\end{table}
\begin{table}
  \begin{tabular}{ | l | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ & Speedup &SM    & Top Stall Reas.\\\hline
    P2G-Atomic		   &    6,933             & -       &82.9\%&S. Scoreb.(24.7\%)\\\hline
    indirect		   &    6,276             & 1.10x   &90.8\%&S.Scoreb.(31.7\%)\\\hline
    batching = 4           &    4,929             & 1.41x   &65.1\%&S. Scoreb.(24.3\%)\\\hline
    both                   &    3,148             & 2.20x   &65.0\%&No Instr.(23.2\%)\\\hline
    \hline
    P2G-Sync               &    3,161             & -       &72.4\%&No Instr.(17.7\%) \\\hline
    batching=4             &    2,595             & 1.21x   &67.0\%&No Instr.(23.7\%) \\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. They form a rotated (unsorted) cube with four particles per cell. Block size is (8,4,4).}\label{table:p2g_optim}
\end{table}

\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM    \\\hline
    global                 &   44,442             &-               &4.6 \%&34.4\%&7.7\%  \\\hline
    \cite{Meyer2015}       &   45,342             &0.98x           &25.1\%&42.4\%&11.7\% \\\hline
    \cite{Meyer2015} sorted&   23,007		  &1.97x           &43.8\%&59.0\%&23.9\% \\\hline
    global sorted          &   20,484		  &2.21x           &7.0 \%&44.0\%&16.1\% \\\hline
    \hline
    P2G-Pull               &    4,747             &9.55x           &3.7\% &6.7\% &39.4\%\\\hline
    P2G-Atomic*		   &    3,148             &14.40x          &5.3\% &6.7\% &65.0\%\\\hline
    P2G-Sync*              &    2,595             &17.47x          &5.9\%&7.6\%  &67.0\%\\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. They form a rotated (unsorted) cube with four particles per cell. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_uniform4}
\end{table}
Barrier 26.3\% on P2G-Sync
\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM    \\\hline
    global                 &  254,410             &-               &17.8\%&9.6\% &1.7\%  \\\hline
    \cite{Meyer2015}       &  217,997             &1.17x           &21.7\%&3.7\% &13.2\% \\\hline
    \cite{Meyer2015} sorted&   30,477		  &8.35x           &64.2\%&64.2\%&22.1\% \\\hline
    global sorted          &   21,413		  &11.88x          &7.2\% &65.8\%&19.4\% \\\hline
    \hline
    P2G-Pull               &   26,822             &9.46x           &1.6\% &2.3\% &35.2\%\\\hline
    P2G-Atomic*		   &   15,552             &16.36x          &3.7\% &7.1\% &65.4\%\\\hline
    P2G-Sync*              &   12,801             &19.87x          &3.6\% &7.3\% &61.3\%\\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million randomly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_random}
\end{table}
\begin{table}
  \begin{tabular}{ | l | r | r | r | r |}    \hline
    Method                 &  Area=47.6\% & Area=23.8\% &Area=11.9\%\\\hline
    P2G-Sync		   &  3,375$\mu s$& 3,266$\mu s$ & 3,129$\mu s$\\\hline
    batching = 2           & -&2,942$\mu s$ & 2,745$\mu s$\\\hline
    batching = 4           & -&- &2,622$\mu s$\\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. They form a rotated (unsorted) cube with one/two/four particles per cell corresponding to the area of the grid that is covered. Block size is (4,4,4).}\label{table:p2g_optim}
\end{table}

>>>>

\newenvironment{code}{\captionsetup{type=algorithm}}{}
\begin{document}
\author{Fabian Meyer}
\title{GPU-Beschleunigung der Material Point Method}
\zweitgutachter{Bastian Krayer, M.Sc.}
\zweitgutachterInfo{(Institut f{\"u}r Computervisualistik, AG Computergraphik)}
% Umschalten der Sprache (fuer englische Rubrikbezeichnungen etc.)

@@

\subsection{Counting Sort \& Stream Compaction}\label{sec:sort}
Sorting on the GPU can dramatically increase workload performance of subsequent steps. If they can profit from following things:
\begin{enumerate}
  \item Accesses can now be handled in a coalesced fashion. \label{it:sort_1}
  \item Data can now be reused due to L2-Cache and shared memory. \label{it:sort_2}
\end{enumerate}
This is the case for algorithms which require neighbor communication.
Supported is a full deep copy of all particle variables. A deep copy profits from item \ref{it:sort_1} and \ref{it:sort_2}, but takes the hit in performance of uncoalesced accesses for all variables. Or, alternatively an indexing to access an existing particle buffer in the specified sorted manner. Indexing only profits from item \ref{it:sort_2} and subsequent shaders will be uncoalesced.

@@

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Method         &  $\Delta t_c$ & Speedup &VRAM              & SM              & Sel. Warp-Stall Reas.\\\hline
    Interleaved    & 305                   & -       &23.0\%          & \textbf{60.9}\%  & S. Scoreb.(17.2\%)\\\hline
    Sequential     & 141                   & 2.16x   &\textbf{49.8}\%   & 37.1\%         & \textbf{S. Scoreb.(2.0\%)} \\\hline
    Seq. (2x)      & 100                   & 3.05x   &69.5\%            & 26.2\%         & L. Scoreb.(80.1\%)\\\hline
    Seq. (128x)    & 98                    & 3.1x    &\textbf{72.9}\%   & 16.9\%         & L. Scoreb.(84.4\%)\\\hline
    Seq. (256x)    & 101                   & 3.0x    &66.4\%            & 14.6\%         & L. Scoreb.(76.9\%)\\
    \hline
  \end{tabular}
  \caption{Optimization of one parallel reduction dispatch on $1024\times1024$ \texttt{vec4} with \texttt{unary\_op(vec4 x)=length(x)} with block\_size = 1024. $\Delta t_c$ is in $\mu s$ Note that the interleaved addressing in this example intentionally commits bank conflicts to show performance impacts. The number in parentheses states the number of additional added sequential elements. Reduced is by block\_size$\times$ sequential elements. Thus they do more work, while also performing better!}\label{tab:map_reduce}
\end{table}

@@

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Interleaved       &  $\Delta t_c$         & Speedup &VRAM              & SM      & L2     & Sel. Stall Reas.\\\hline
    Conflicts           & 748                   & -       &17.5\%            & 51.7\%  & 8.2\%  & S. Scoreb.(6.3\%)\\\hline
    No confl.           & 571                   & 1.31x   &25.7\%            & 50.1\%  & 5.4\%  & \textbf{S. Scoreb.(2.6\%)}\\\hline
    part. confl.        & 546                   & 1.36x   &23.5\%            & 50.8\%  & 6.0\%  & \textbf{S. Scoreb.(3.2\%)} \\\hline\hline
    Seq. (2x)           & 323                   & 2.31x   &41.7\%            & \textbf{43.0}\%  & 17.2\% & S. Scoreb.(32.0\%)\\\hline
    unrolled            & 258                   & 2.90x   &\textbf{53.8}\%   & 42.2\%  & \textbf{22.0}\% & S. Scoreb.(32.0\%)\\\hline\hline
    Seq. (4x)           & 311                   & 2.40x   &56.8\%            & 22.7\%  & 37.9\% & L. Scoreb.(43.3\%)\\\hline
    unrolled            & 297                   & 2.52x   &61.1\%            & 21.2\%  & 40.6\% & L. Scoreb.(49.9\%)\\
    \hline
  \end{tabular}
  \caption{Optimization of one scan dispatch on $4 \times 1024\times1024$ integers with \texttt{binary\_op(x,y) := x+y}. $\Delta t_c$ in $\mu s$, block size is 1024. The number in parentheses states the number of additionally loaded sequential elements. Segment of each scan is block\_size$\times$ sequential elements. Higher reduction factor corresponds to more work being done.}\label{tab:scan}
\end{table}
The second optimization is giving these algorithms more sequential work. This will lighten the SM throughput and increase VRAM throughput as can be seen in both cases. Note, that each thread group in table \ref{tab:map_reduce} and \ref{tab:scan} does the same amount of parallel work; the sequential work is additionally being done by each thread. However due to the additional sequential work parallelization the amount of thread groups go down by the same factor. As mentioned in \cite{NVIDIA:PARALLEL_REDUCTION} this saturates the threads and thus improves performance.

@@

For the scan this is a little bit harder as elements need to be stored in registers. The effects can be seen in table \ref{tab:scan}. The segment each scan can handle is by far not as extensive as in parallel reduction. But the speedup achieved from a partially conflict free to an implementation with sequential work is in contrast much higher. Since each thread requires the nearest by elements, L2-cache hits increase rapidly as does the L2 utilization. Thus the scoreboard stall reasons rise again. Unrolling the static loop which loads the sequential elements additionally hides memory latency.

\textbf{Counting Sort:} The aim of sorting is coalesced as well as increased caching behavior for all following shaders. As a motivation the binning's similarity to the MPM-Operations might already be highlighted. A key difference is that binning does not query neighbor information (in the global atomic approach). Table \ref{table:bin} shows how a sorted particle position increases coalescing and L2-Cache hits. The VRAM-throughput increases to 75.3\% due to coalescing. Nearby particles hit the same bins and thus L2 rises up to 35.0\%. The 6.95x speed up shows how important the sorting as a preprocess is.
\begin{table}[htbp]
  \begin{tabular}{ | l | r | r | r | r | r | r |}    \hline
    Ordering            &  $\Delta t_c (\mu s)$ & Speedup &VRAM              & SM      & L2     & L2-Hit \\\hline
    Random              &  1,516                & -       &25.0\%            & 3.4\%   & 9.1\%  & 10.8\% \\\hline
    Deep sort           &    218                & 6.95x   &\textbf{75.3\%}   & 24.4\%  & \textbf{35.0\%} & \textbf{37.8\%} \\
    \hline
 \end{tabular}
\caption{Order dependency of binning of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid.}\label{table:bin}
\end{table}

Table \ref{table:sort} shows the performance of index and deep sorting as is. The index sorting will have constant timing across cases. Their writing access will stay uncoalesced. Moreover shaders down the pipeline will also not benefit from coalescing. The binning in table \ref{table:bin} thus for instance always runs on random ordering. Global atomic binning does not use neighbor data. Thus index sorting is not affecting it.

Again differences similarly to the binning can be seen when already working with a sorted state. Coalescing highly increases VRAM throughput. So much that a deep sort on eight \texttt{vec4}s is almost on equal footing with a random index sort. The speedup from an uncoalesced to a fully coalesced state is 5.68x. The full sort suffers for each (particle) variable from an uncoalesced access. However if sorting is not done the MPM-Transfers will then suffer from those.

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r| }    \hline
    Best Case    &  $\Delta t_c (\mu s)$&VRAM  & Top Stall Reas.    & Read   & Write  \\\hline
    Index        &  1,516       &25.6\%& L. Scoreb.(78.7\%) & 16.30\%& 9.30\% \\\hline
    Full(8 vec4) &  1,731       &\textbf{75.3}\%& L. Scoreb.(92.2\%) & 38.60\%& 34.60\%\\
    \hline
    \hline
    Worst Case             &  $\Delta t_c(\mu s)$         &VRAM              & Top Stall Reas.    & Read   & Write\\\hline
    Index                  &  1,516                &26.0\%            & L. Scoreb.(73.5\%) & 16.10\%& 9.40\%\\\hline
    Full(8 vec4)           &  9,847                &24.0\%            & L. Scoreb.(82.2\%) & 9.90\% & 14.0\%\\
    \hline
 \end{tabular}
\caption{Order dependency of reordering of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid. Best case is already sorted, worst case is unsorted and still fully random.}\label{table:sort}
\end{table}

Binning as well as reordering will rarely run fully sorted as mentioned in \ref{sec:sort}. Sorting works on the last sorted state. The CFL-condition encourages limited reach. The Material-Point-Method encourages uniform movement. These two properties in general will guarantee higher coalescing as well as L2-cache hit rates. It is however up to the dynamics of the problem how much time is spend sorting.

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r |r|r|r|}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2    &SM     \\\hline
    G2P global             &   38,318             &-               &24.3\%&34.4\% & 9.6\% \\\hline
    sorted                 &    2,991		  &12.81x          & 7.5\%&52.3\% &67.1\% \\\hline
    G2P-Shared             &    9,665             & 3.96x          & 3.4\%& 2.5\% &65.0\% \\\hline
 \end{tabular}
 \caption{G2P-Transfers of a $128\times 128\times128$ grid back to one million randomly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$. Block size is (8,4,4).}\label{table:p2g_uniform4}
\end{table}

VRAM-L.Scoreboard 79.9\%
Wait No Instructions
\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM    \\\hline
    G2P global             &   10,316             &-               &2.3\%&34.4\%&32.4\% \\\hline
    sorted                 &    2,629		  &3.92x           &4.1\%&9.4\%&66.9\%  \\\hline
    G2P-Shared             &    2,232             &4.62x           &7.3\%&4.1\% &63.6\% \\\hline
 \end{tabular}

 \caption{G2P-Transfers of a $128\times 128\times128$ grid back to one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$. They form a rotated (unsorted) cube with four particles per cell. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_uniform4}
\end{table}
\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ & Speedup &SM    & Top Stall Reas.\\\hline
    P2G-Atomic		   &    6,933             & -       &82.9\%&S. Scoreb.(24.7\%)\\\hline
    indirect		   &    6,276             & 1.10x   &90.8\%&S.Scoreb.(31.7\%)\\\hline
    batching = 4           &    4,929             & 1.41x   &65.1\%&S. Scoreb.(24.3\%)\\\hline
    both                   &    3,148             & 2.20x   &65.0\%&No Instr.(23.2\%)\\\hline
    \hline
    P2G-Sync               &    3,161             & -       &72.4\%&No Instr.(17.7\%) \\\hline
    batching=4             &    2,595             & 1.21x   &67.0\%&No Instr.(23.7\%) \\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. They form a rotated (unsorted) cube with four particles per cell. Block size is (8,4,4).}\label{table:p2g_optim}
\end{table}

\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM    \\\hline
    global                 &   44,442             &-               &4.6 \%&34.4\%&7.7\%  \\\hline
    \cite{Meyer2015}       &   45,342             &0.98x           &25.1\%&42.4\%&11.7\% \\\hline
    \cite{Meyer2015} sorted&   23,007		  &1.97x           &43.8\%&59.0\%&23.9\% \\\hline
    global sorted          &   20,484		  &2.21x           &7.0 \%&44.0\%&16.1\% \\\hline
    \hline
    P2G-Pull               &    4,747             &9.55x           &3.7\% &6.7\% &39.4\%\\\hline
    P2G-Atomic*		   &    3,148             &14.40x          &5.3\% &6.7\% &65.0\%\\\hline
    P2G-Sync*              &    2,595             &17.47x          &5.9\%&7.6\%  &67.0\%\\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. They form a rotated (unsorted) cube with four particles per cell. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_uniform4}
\end{table}
Barrier 26.3\% on P2G-Sync
\begin{table}[htpb]
  \begin{tabular}{ | l | r | r | r | r | r |}    \hline
    Method                 &  $\Delta t_c(\mu s)$ &Speedup         &VRAM  & L2   &SM    \\\hline
    global                 &  254,410             &-               &17.8\%&9.6\% &1.7\%  \\\hline
    \cite{Meyer2015}       &  217,997             &1.17x           &21.7\%&3.7\% &13.2\% \\\hline
    \cite{Meyer2015} sorted&   30,477		  &8.35x           &64.2\%&64.2\%&22.1\% \\\hline
    global sorted          &   21,413		  &11.88x          &7.2\% &65.8\%&19.4\% \\\hline
    \hline
    P2G-Pull               &   26,822             &9.46x           &1.6\% &2.3\% &35.2\%\\\hline
    P2G-Atomic*		   &   15,552             &16.36x          &3.7\% &7.1\% &65.4\%\\\hline
    P2G-Sync*              &   12,801             &19.87x          &3.6\% &7.3\% &61.3\%\\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million randomly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. Block size is (8,4,4). Methods marked with a star(*) are executed with batching = 4.}\label{table:p2g_random}
\end{table}
\begin{table}
  \begin{tabular}{ | l | r | r | r | r |}    \hline
    Method                 &  Area=47.6\% & Area=23.8\% &Area=11.9\%\\\hline
    P2G-Sync		   &  3,375$\mu s$& 3,266$\mu s$ & 3,129$\mu s$\\\hline
    batching = 2           & -&2,942$\mu s$ & 2,745$\mu s$\\\hline
    batching = 4           & -&- &2,622$\mu s$\\\hline
 \end{tabular}
 \caption{P2G-Transfers of one million uniformly positioned particles with random velocities between $v_x,v_y,v_z \in [-1.0;1.0]$ in a $128\times 128\times128$ grid. They form a rotated (unsorted) cube with one/two/four particles per cell corresponding to the area of the grid that is covered. Block size is (4,4,4).}\label{table:p2g_optim}
\end{table}
